### 2023-05-28 paper/Chapters/Chapter4.tex
WIP
Update the paper and the evaluation

@@

@@

We chose to minimize the reprojection error between the board and back-projected
corners, which were initially detected. The projection and back-projection are
the inverse of each other, hence minimizing the error between the projection of
the board and the detected corners and minimizing the error between the
back-projection of the detected corners and the board are
equivalent\todo{Doesn't sound very rigorous to me.}.

@@

@@

@@

Given the possible positions of the previously undetected corners, we used a
classifier, proposed by \cite{geigerAutomaticCameraRange2012}.
By convolving the image with the templates of the board corners\todo{I have to
reference it from the related work section}, we obtained the
probability\todo{It's not probability} of
each pixel being a corner.\todo{Also, explain in details + math how it's done}

>>>>

For the feature detection, we used the approach proposed by
\cite{geigerAutomaticCameraRange2012} (see \cref{sec:boards_features_detection,
	feature_detection}).
\missingfigure{Show the feature detection results.}

@@

In this work, we obtained the camera calibration in two steps: first, we
used the approach, proposed by \cite{scaramuzzaToolboxEasilyCalibrating2006} to
obtain the initial approximation of the camera parameters. Then, we used the
optimization to minimize the reprojection error between the board and the
back-projected corners.

@@

We chose to minimize the reprojection error between the board and back-projected
corners, which were initially detected. The projection and back-projection are
the inverse of each other, hence minimizing the error between the projection of
the board and the detected corners and minimizing the error between the
back-projection of the detected corners and the board are
equivalent\todo{Doesn't sound very rigorous to me.}. We minimized the error
between the board and the back-projected corners because back-projection does
not require the root finding.

@@

\subsection{Initial approximation}\label{sub:initial_approximation}

\cite{scaramuzzaToolboxEasilyCalibrating2006} proposed an automatic method for
camera calibration, which consisted of the following steps:
\begin{itemize}
	\item Solving for the camera extrinsic parameters
	\item Solving for the camera intrinsic parameters
	\item Linear refinement of the intrinsic and extrinsic parameters
	\item Iterative center detection
	\item Non-linear refinement
\end{itemize}

In this work, we focused on the single image, while the original approach relied
on the multiple images. Therefore, we used only the first two steps of the
algorithm.

\subsubsection{Solving for the camera extrinsic parameters}\label{ssub:solving_for_the_camera_extrinsic_parameters}

To derive the solver for the camera extrinsic parameters, start from
\cref{eq:projection}:
\begin{align}
	\alpha \mathbf{u}                                  & = K f(H\mathbf{X})                  \\
	\alpha K^{-1} \mathbf{u}                           & = f(H\mathbf{X})    &  &
	\text{Move \(K\) to the left side}                                                       \\
	\alpha g( K^{-1}\mathbf{u})                        & = g(f(H\mathbf{X})) &  &
	\text{Set \(\mathbf{\widehat{u}} = K^{-1}\mathbf{u}\); apply \(g(\cdot)\) to both sides} \\
	\alpha \begin{pmatrix}
		       \widehat{u}_x \\ \widehat{u}_y \\ \psi(r(\mathbf{\widehat{u}}))
	       \end{pmatrix}^{T} & = H\mathbf{X}                                 \\
\end{align}

For \(K\) we used the placeholder values. \(f\) was set to a constant value for
the typical consumer camera, and \(c_x, c_y\) were set to the center of the
image.
The correct values will be computed during the optimization \cref{sub:optimization}.

To eliminate the dependency on the scale \(\alpha\), multiply both sides
vectorially by \(g(\mathbf{\widehat{u}})\):
\begin{equation}
	\alpha g(\mathbf{\widehat{u}}) \times g(\mathbf{\widehat{u}})
	= g(\mathbf{\widehat{u}}) \times H\mathbf{X}
	= 0 \implies
	\begin{pmatrix}
		\widehat{u}, \widehat{v}, \psi(r(\mathbf{\widehat{u}}))
	\end{pmatrix}^{T} \times \begin{bmatrix}
		\mathbf{r_1} & \mathbf{r_2} & \mathbf{t}
	\end{bmatrix}^{T} \mathbf{X} = 0
	\label{eq:eq_no_multiplier}
\end{equation}.

From \cref{eq:eq_no_multiplier}, we can see that a point contributes to three
homogeneous equations:
\begin{align}
	\widehat{v} (t_1X + t_2 Y + t_3) -
	g(r(\mathbf{\widehat{u}})) (r_{12}X + r_{22}Y + t_2 ) & = 0
	\label{eq:scaramuzza_system_1}                              \\
	g(r(\mathbf{\widehat{u}})) (r_{11}X + r_{12}Y + t_1) -
	\widehat{u} (t_1X + t_2 Y + t_3)                      & = 0
	\label{eq:scaramuzza_system_2}                              \\
	\widehat{u} (r_{12}X + r_{22}Y + t_2 ) -
	\widehat{v} (r_{11}X + r_{12}Y + t_1)                 & = 0
	\label{eq:scaramuzza_system_3}
\end{align}

Only \cref{eq:scaramuzza_system_3} is linear in the unknowns. Each point gives a
single equation. Now, by rewriting the equation in the matrix form
\(M \cdot H = 0\), where \(H = \begin{pmatrix}
	r_{11}, r_{12}, r_{12}, r_{22}, t_1, t_2
\end{pmatrix}\) we get:

\begin{equation}
	M = \begin{bmatrix}
		-\widehat{v}_1 X_1 & -\widehat{v}_1 Y_1 & -\widehat{u}_1 X_1 & -\widehat{u}_1 Y_1 & -\widehat{v}_1 & -\widehat{u}_1 \\
		\vdots             & \vdots             & \vdots             & \vdots             & \vdots         & \vdots         \\
		-\widehat{v}_N X_N & -\widehat{v}_N Y_N & -\widehat{u}_N X_N & -\widehat{u}_N Y_N & -\widehat{v}_N & -\widehat{u}_N
	\end{bmatrix}
\end{equation}, where \(N\) is the number of points.

The linear estimate of \(H\) is found by minimizing \(\left\lVert M \cdot H
\right\rVert ^{2}\) using SVD. The solution is known up to a scale factor.

To find \(t_1\) and \(t_2 \), note that \(\mathbf{r_1}\) and
\(\mathbf{r_2}\) are orthonormal:
\begin{subnumcases}{}
	\lambda^{2} r_{11} r_{12} + \lambda^{2} r_{12} r_{22} + \lambda^{2} t_1
	t_2  = 0
	\label{eq:orthonormality_1}                                                     \\
	\lambda \sqrt{r_{11}^{2} + r_{12}^{2} + t_1^{2}}                            = 1
	\label{eq:orthonormality_2}                                                     \\
	\lambda \sqrt{r_{12}^{2} + r_{22}^{2} + t_2 ^{2}}                            = 1
	\label{eq:orthonormality_3}
\end{subnumcases},
where \(\lambda\) is non-zero multiplier.

Now, to solve for \(t_1\) and \(t_2 \), first find possible values for
\(t_2 ^{2}\):
\begin{align}
	- \frac{r_{11} r_{12} + r_{12} r_{22}}{t_2 }                  & = t_1 &  &
	\text{Solve eq. (\ref{eq:orthonormality_1}) for \(t_1\)}
	\label{eq:orthonormality_4}                                                \\
	\\
	\sqrt{\lambda^{2} \left( \frac{r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2}\right)}                               & = 1   &  &
	\text{Substitute into eq. (\ref{eq:orthonormality_2})}                     \\
	\lambda^{2} (\frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2})                                      & = 1   &  &
	\text{Square both sides}                                                   \\
	\lambda^{2} (\frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2} - r_{12}^{2} - r_{22}^{2} - t_2 ^{2}) & = 0   &  &
	\text{Subtract eq. (\ref{eq:orthonormality_3})}
	\label{eq:orthonormality_5}
	\\
	\frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2} - r_{12}^{2} - r_{22}^{2} - t_2 ^{2}  & = 0   &  &
	\text{Divide both by \(\lambda ^{2}\)}                                     \\
	t_2 ^{4} - (r_{11}^{2} + r_{12}^{2} - r_{12}^{2} -
	r_{22}^{2}) t_2 ^{2} - (r_{11} r_{12} + r_{12} r_{22})^{2}    & = 0   &  &
	\text{Multiply by \(t_2 ^{2}\)}
	\label{eq:orthonormality_6}
\end{align}

Now, solve \cref{eq:orthonormality_6} for \(t_2 ^{2}\), and take a root to find
possible values for \(t_2 \).

To find \(t_1\), substitute the found values for \(t_2 \) into
\cref{eq:orthonormality_4} or \cref{eq:orthonormality_5} depending on the value
of \(t_2 \):
\begin{subnumcases}{}
	t_1  = - \frac{r_{11} r_{12} + r_{12} r_{22}}{t_2 } &
	\text{\(t_2  \neq 0\)} \\
	t_1  =  r_{11}^{2} + r_{12}^{2} - r_{12}^{2} - r_{22}^{2} &
	\text{\(t_2  = 0\)}
\end{subnumcases}

Now, it's possible to find \(H\) for each of the pairs of \(t_1\) and
\(t_2 \).

Lastly, to select the correct \(H\), the author assumes that one of the boards'
corners has the coordinates \( \begin{pmatrix}
	0, 0
\end{pmatrix}^{T}\). Then, the rotation wouldn't affect this
point, and it would be projected to   \(\begin{pmatrix}
	t_1, t_2
\end{pmatrix}^{T}\). Hence, the best matrix would be such that has the closest
\(\begin{pmatrix}
	t_1, t_2
\end{pmatrix}^{T}\) to \(\begin{pmatrix}
	X, Y
\end{pmatrix}^{T}\) of the corner, which is associated with the board's corner
with corrdinates \(\begin{pmatrix}
	0, 0
\end{pmatrix}^{T}\).

However, often the algorithm finds who matrices \(H\) such that they have the
same \(t_1\) and \(t_2 \), but different \(r_{11}\), \(r_{12}\), \(r_{21}\). To
find the best matrix, we found the intrinsic values for each of them, and
backprojected the board's corners using both matrices.
Then, we used the one which gave the smallest reprojection error.

\subsubsection{Solving for the camera intrinsic parameters}\label{subsub:solving_for_the_camera_intrinsic_parameters}

Now, to find the rest of the parameters, we substitute the values, found in the
previous step into \cref{eq:scaramuzza_system_1} and
\cref{eq:scaramuzza_system_2}. We assumed the number of the division model's
parameter to be equal to 2, and the scalar multiplier to be equal to 1
\cref{subsub:back_projection_using_the_division_model}:

\begin{equation}
	\begin{bmatrix}
		A_1 \rho_1^{2} & A_1 \rho_1^{4} & -v_1   \\
		C_1 \rho_1^{2} & C_1 \rho_1^{4} & -v_1   \\
		\cdots         & \cdots         & \cdots \\
		A_N \rho_N^{2} & A_N \rho_N^{4} & -v_N   \\
		C_N \rho_N^{2} & C_N \rho_N^{4} & -v_N
	\end{bmatrix} \cdot \begin{bmatrix}
		\lambda_1 \\
		\lambda_2 \\
		t_3
	\end{bmatrix} = \begin{bmatrix}
		B_1 - A_1 \\
		D_1 - C_1 \\
		\cdots    \\
		B_N - A_N \\
		D_N - C_N
	\end{bmatrix}
\end{equation},
where
\begin{align}
	A_i & = r_{21} X_i + r_{22} Y_i + t_2 \\
	B_i & = v_i (r_{31} X_i + r_{32} Y_i) \\
	C_i & = r_{11} X_i + r_{12} Y_i + t_1 \\
	D_i & = u_i (r_{31} X_i + r_{32} Y_i) \\
\end{align}

The solution can be found using the least squares method.

@@

\section{Additional features detection}\label{sec:additional_features_detection}

@@

In this work, we used the method, proposed by
\cite{geigerAutomaticCameraRange2012} as it didn't require the whole board to be
visible, automatically determines the board's number of rows and columns and
worked quite well on highly-distorted images.

In this work, we directly used the following steps from the algorithm:
\begin{enumerate}
	\item Corner detection
	\item Sub-pixel corner and orientation refinement
\end{enumerate}

\subsection{Corner detection}\label{sub:corner_detection}

% \begin{multicols}{2}
%
% 	According to \cite{geigerAutomaticCameraRange2012}, the following approach proved to be
% 	more robust to image clutter and blur than other common choices
% 	\citep{harrisCombinedCornerEdge1988, shiGoodFeaturesTrack2000}.
%
% 	To detect corners in a grayscale image $I$, the author used two $n \times n$ prototypes
% 	for axis-aligned and $45^{\circ}$ rotated corners, respectively.
% 	Each prototype
% 	is constructed using four filter kernels $\{A, B, C, D\}$. The corner likelihood
% 	$c$ at each pixel is computed by:
%
% 	\columnbreak
%
% 	\begin{figure}[H]
% 		\centering
%
% 		\begin{subfigure}{0.48\textwidth}
% 			\centering
% 			\includegraphics[width=0.23\textwidth]{./template_1_a.png}
% 			\includegraphics[width=0.23\textwidth]{./template_1_b.png}
% 			\includegraphics[width=0.23\textwidth]{./template_1_c.png}
% 			\includegraphics[width=0.23\textwidth]{./template_1_d.png}
% 			\caption{Corner prototype 1 (A, B, C, D)}
% 		\end{subfigure}
%
%     \vspace{0.5cm}
%
% 		\begin{subfigure}{0.48\textwidth}
% 			\centering
% 			\includegraphics[width=0.23\textwidth]{./template_2_a.png}
% 			\includegraphics[width=0.23\textwidth]{./template_2_b.png}
% 			\includegraphics[width=0.23\textwidth]{./template_2_c.png}
% 			\includegraphics[width=0.23\textwidth]{./template_2_d.png}
% 			\caption{Corner prototype 2 (A, B, C, D)}
% 		\end{subfigure}
%
% 		\caption{Corner prototypes}
% 	\end{figure}
%
% \end{multicols}

\begin{figure}[h]
	\centering

	\begin{subfigure}{0.45\textwidth}
		\centering
		\begin{minipage}{\textwidth}
			\includegraphics[width=0.24\textwidth]{./template_1_a.png}
			\includegraphics[width=0.24\textwidth]{./template_1_b.png}
			\includegraphics[width=0.24\textwidth]{./template_1_c.png}
			\includegraphics[width=0.24\textwidth]{./template_1_d.png}
		\end{minipage}
		\caption{Corner prototype 1 (A, B, C, D)}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.45\textwidth}
		\centering
		\begin{minipage}{\textwidth}
			\includegraphics[width=0.24\textwidth]{./template_2_a.png}
			\includegraphics[width=0.24\textwidth]{./template_2_b.png}
			\includegraphics[width=0.24\textwidth]{./template_2_c.png}
			\includegraphics[width=0.24\textwidth]{./template_2_d.png}
		\end{minipage}
		\caption{Corner prototype 2 (A, B, C, D)}
	\end{subfigure}

	\caption{Corner prototypes \citep{geigerAutomaticCameraRange2012}}
	\label{fig:corner_prototypes}
\end{figure}

According to \cite{geigerAutomaticCameraRange2012}, the following approach proved to be
more robust to image clutter and blur than other common choices
\citep{harrisCombinedCornerEdge1988, shiGoodFeaturesTrack2000}.

To detect corners in a grayscale image $I$, the author used two $n \times n$ prototypes
for axis-aligned and $45^{\circ}$ rotated corners, respectively.
Each prototype
is constructed using four filter kernels $\{A, B, C, D\}$ (\cref{fig:corner_prototypes}). The corner likelihood
$c$ at each pixel is computed by:

\begin{equation}
	\begin{aligned}
		c     & =\max \left(s_1^1, s_2^1, s_1^2, s_2^2\right)                                             \\
		s_1^i & =\min \left(\min \left(f_A^i, f_B^i\right)-\mu, \mu-\min \left(f_C^i, f_D^i\right)\right) \\
		s_2^i & =\min \left(\mu-\min \left(f_A^i, f_B^i\right), \min \left(f_C^i, f_D^i\right)-\mu\right) \\
		\mu   & =0.25\left(f_A^i+f_B^i+f_C^i+f_D^i\right)
	\end{aligned}
\end{equation}

Then, the authors apply the conservative non-maximum suppression to the
response map, and additionally filter the corners by assuring that there are two
modes of the gradient directions.

\subsection{Sub-pixel Corner and Orientation Refinement}\label{sub:sub_pixel_corner_and_orientation_refinement}

Sub-pixel accurate corner locations and edge orientations are refined using the following optimizations:

For sub-pixel corner localization, the corner $\mathbf{c}$ in a 2D space
$\mathbb{R}^2$ is refined using the assumption that the gradient direction
\(\mathbf{g_{p}}\) at
each pixel  \(\mathbf{p}\) in the small neighbourhood around the corner is
orthogonal to \(\mathbf{p} - \mathbf{c}\).

\begin{equation}
	\mathbf{c}=\arg \min _{\mathbf{c}^{\prime}} \sum_{\mathbf{p} \in
	\mathcal{N}_{\mathbf{I}}\left(\mathbf{c}^{\prime}\right)}\left(\mathbf{g}_{\mathbf{p}}^T\left(\mathbf{p}-\mathbf{c}^{\prime}\right)\right)^2
\end{equation}

For the orientation refinement, authors minimize the distance from the
orientation vectors \(\mathbf{e_1}, \mathbf{e_2}\) to the gradient directions
\(\mathbf{g_p}\) in the neighbourhood of the corner.

\begin{equation}
	\mathbf{e}_i=\arg \min _{\mathbf{e}_i^{\prime}} \sum_{\mathbf{p} \in
		\mathcal{M}_i}\left(\mathbf{g}_{\mathbf{p}}^T \mathbf{e}_i^{\prime}\right)^2
	\quad \text { s.t. } \mathbf{e}_i^{\prime T} \mathbf{e}_i^{\prime}=1
\end{equation}.
